RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-09-15 doc/paper.tex
Intro: working...

	 There are many different ways to tackle the SR by achieving the right analytic expressions, most of them are related to machine learning techniques or to the technique of genetic programming (GP) and various approximate methods, such as meta-heuristics. Among the first GP methods to tackle SR is the one of Raidl~\cite{raidl1998hybrid} based on a hybrid variant of genetic programming. A differential evolution algorithm was proposed by Cerny et al.~\cite{cerny2008using}. Age-fitness Pareto Optimization approach is proposed by Smidt and Lipson~\cite{schmidt2010age}.  
	 Application of local--based heuristics on solving SR is discusses by Commenda in his Ph.D. thesis~\cite{kommenda2018local}. A  GP-based approach, the gene-pool optimal mixing evolutionary algorithm (GOMEA) is studied by Virgolin et al.~\cite{virgolin2021improving}.  Another evolutionary algorithm, the interaction-transformation EA (ITEA) has been proposed by de Franca et al.~\cite{de2021interaction}. Simulated annealing to solve SR is proposed by Kantor~\cite{kantor2021simulated}. Kommenda et al.~\cite{kommenda2020parameter} proposed a method called OPERON algorithm, which uses nonlinear least squares for parameter
	 identification of SR models further integrated into a local search mechanism in
	 tree-based GP. The C++ implementation of OPERON is discussed in~\cite{burlacu2020operon}. The metod that uses Taylor polynomial to approximate the symbolic equation that fits the dataset, called Taylor genetic programming, is proposed in~\cite{he2022taylor}. A GP approach that uses the idea of semantic back-propagation (SBP-GP) is proposed in~\cite{virgolin2019linear}.   Empirical analysis of variance between many GP--based methods for SR is discussed in~\cite{kammerer2021empirical}. It is also worth to mention the GP--based Eureqa commercial solver~\cite{schmidt2009distilling, schmidt2011machine} that uses age-fitness pareto optimization  with co-evolved fitness estimation. This solver is nowadays accepted for the gold standard of symbolic regression.

@@

	Powerful hybrid techniques to solve SR are also well studied; among them, we emphasize the EPLEX solver from~\cite{la2019probabilistic,la2016epsilon} and AI Feynman algorithm from~\cite{udrescu2020ai}, a physics-inspired divide-and-conquer method that  combines neural network
	fitting. The later one is one of the most efficient method for physically inspired models. We also mention the Fast Function Extraction (FFX) algorithm developed by McConaghy~\cite{mcconaghy2011ffx}, which is a non-evolutionary technique in combination with a machine learning technique --- pathwise regularized learning -- which is in charge to quickly prune a huge set of candidate basis functions down to compact models.

	Short overview of the GP--based methods to solve SR are given in Table~\ref{tab:gp-based} \fxnote{TODO}

	 \begin{table}[htb]
		\centering
		\begin{tabularx}{350pt}{l  l  X}  
			Algorithm          &   paper/year &   Short details   \\ \hline
	        Hybrid GP  &   \cite{raidl1998hybrid} (1998)   & GP--based method; solutions locally optimizes by method of least squares to find optimum coefficients for top-level terms \\ 
	        DFE        & \cite{cerny2008using} (2008)   & Differential evolution algorithm \\
			Eurequa                & \cite{schmidt2009distilling, schmidt2011machine} (2009, 2011) & Age-fitness pareto optimization approach using co-evolved fitness estimation   \\
			APF            &      \cite{schmidt2010age} (2010)        &  Age-fitness pareto optimization approach                        \\
			EPLEX          &      \cite{la2016epsilon} (2016)         &   A parent selection method called $\epsilon$--lexicase selection                \\ 
	       Local optimization     & \cite{kommenda2018local}   & Constants Optimization in GP by Nonlinear Least Square \\
	       \fxnote{TODO}             &      & 
		\end{tabularx}
		\caption{Overview of methods to solve SR.}
		\label{tab:gp-based}
	\end{table}

		Short overview of the machine--learning based method to solve SR are given in Table~\ref{} \fxnote{TODO}

		\fxnote{TODO: about benchmarking in the literature...}
\subsection{Contributions of this work}

>>>>

	 There are many different ways to tackle the SR by achieving the right analytic expressions, most of them are related to machine learning techniques or to the technique of genetic programming (GP) and various approximate methods, such as meta-heuristics. Among the first GP methods to tackle SR is the one of Raidl~\cite{raidl1998hybrid} based on a hybrid variant of genetic programming. A differential evolution algorithm was proposed by Cerny et al.~\cite{cerny2008using}. Age-fitness Pareto Optimization approach is proposed by Smidt and Lipson~\cite{schmidt2010age}.  Application of artificial bee colony programming to solve SR is proposed by Karaboga et al.~\cite{karaboga2012artificial}. 
	 Influence of local--based heuristics on solving SR is reported by Commenda in his Ph.D. thesis~\cite{kommenda2018local}. A  GP-based approach, the gene-pool optimal mixing evolutionary algorithm (GOMEA) is studied by Virgolin et al.~\cite{virgolin2021improving}.  Another evolutionary algorithm, the interaction-transformation EA (ITEA) has been proposed by de Franca et al.~\cite{de2021interaction}. Simulated annealing to solve SR is proposed by Kantor~\cite{kantor2021simulated}. A Variable neighborhood programming approach to solve SR  is proposed by Elleurich et al.~\cite{elleuch2020variable}; this technique is initially proposed in~\cite{elleuch2016variable}. 
	 Kommenda et al.~\cite{kommenda2020parameter} proposed a method called OPERON algorithm, which uses nonlinear least squares for parameter identification of SR models further integrated into a local search mechanism in
	 tree-based GP. The C++ implementation of OPERON is discussed in~\cite{burlacu2020operon}. The method that uses Taylor polynomial to approximate the symbolic equation that fits the dataset, called Taylor genetic programming, is proposed in~\cite{he2022taylor}. A GP approach that uses the idea of semantic back-propagation (SBP-GP) is proposed in~\cite{virgolin2019linear}.   Empirical analysis of variance between many GP--based methods for SR is discussed in~\cite{kammerer2021empirical}. It is also worth to mention the GP--based Eurequa commercial solver~\cite{schmidt2009distilling, schmidt2011machine} that uses age-fitness pareto optimization  with co-evolved fitness estimation. This solver is nowadays accepted for the gold standard of symbolic regression.

@@

	Powerful hybrid techniques to solve SR are also well studied; among them, we emphasize the EPLEX solver from~\cite{la2019probabilistic,la2016epsilon} and AI Feynman algorithm from~\cite{udrescu2020ai}, a physics-inspired divide-and-conquer method that  combines neural network
	fitting. The later one is one of the most efficient method for physically inspired models. We also mention the Fast Function Extraction (FFX) algorithm developed by McConaghy~\cite{mcconaghy2011ffx},  a non-evolutionary method in combination with a machine learning technique---path-wise regularized learning--- which   quickly prune a huge set of candidate basis functions down to compact models.

	A short chronological overview of the most important  literature methods to solve SR are given in Table~\ref{tab:gp-based}.

	 \begin{table}[!ht]
		\centering
		\scalebox{0.7}{
		\begin{tabularx}{450pt}{l  l  X}  
			Algorithm          &   Paper/year &   Short details   \\ \hline
	        Hybrid GP  &   \cite{raidl1998hybrid} (1998)   & GP--based method; solutions locally optimizes by method of least squares to find optimum coefficients for top-level terms \\ 
	        DFE        & \cite{cerny2008using} (2008)   & Differential evolution algorithm \\
	        Eurequa                & \cite{schmidt2009distilling, schmidt2011machine} (2009, 2011) & Age-fitness pareto optimization approach using co-evolved fitness estimation   \\
        	APF            &      \cite{schmidt2010age} (2010)        &  Age-fitness pareto optimization approach                        \\
        	FFX  & \cite{mcconaghy2011ffx} (2011)  & The Fast Function Extraction algorithm --  non-evolutionary technique based on a machine learning technique called path-wise regularized learning \\
             ABCP &  \cite{karaboga2012artificial} (2012)  & Artificial bee colony programming  approach \\
			EPLEX          &      \cite{la2016epsilon} (2016)         &   A parent selection method called $\epsilon$--lexicase selection  \\ 
	       Local optimization     & \cite{kommenda2018local} (2018)   & Constants Optimization in GP by Nonlinear Least Square \\
	       SBP-GP &  \cite{virgolin2019linear} (2019)  & the idea of semantic back-propagation \\
	       BSR & \cite{jin2019bayesian} (2019) & ML--based approach; Bayesian symbolic regression  \\ 
	       DSR  & \cite{petersen2019deep} (2019) & Deep Symbolic Regression based on a RNN approach further utilizing the policy gradient search \\
	       OPERON & \cite{kommenda2020parameter} (2020) &  Utilizing nonlinear least squares for parameter identification of SR models with LS \\
	       VNP    & \cite{elleuch2020variable} (2020) & VNS based GP approach \\
	       OccamNet & \cite{costa2020fast} (2020) &   A fast neural network approach; the model defines a probability distribution over a non-differentiable function space; it samples functions and updates the weights with back-propagation  based on cross-entropy matching in an EA strategy	 \\

	       AI Feynman  & \cite{udrescu2020ai} (2020) & A physics-inspired divide-and-conquer method; it  combines neural network fitting \\
	             GOMEA  & \cite{virgolin2021improving} (2021)    & A model-based
	       EA framework called Gene-pool optimal mixing evolutionary algorithm \\
	       ITEA & \cite{de2021interaction} (2021)   & EA based approach called the interaction-transformation EA   \\
	       SA & \cite{kantor2021simulated} (2021) &  Simulated annealing approach \\

         DRLA & \cite{mundhenk2021symbolic} (2021)  &    A deep reinforcement learning approach enhanced with genetic programming \\
            Taylor GP &  \cite{he2022taylor} (2022)  &  Taylor polynomials approximations  \\ \hline

	\end{tabularx} }
		\caption{Overview of methods to solve SR.}
		\label{tab:gp-based}
	\end{table}

	 \fxnote{TODO: about benchmarking in the literature...}
	 Concerning 
\subsection{Contributions of this work}
